---
title: "Review"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
```

## Instructions

1.  **Clone the Homework Repository**\
    Clone the GitHub repository containing the homework file to your **local machine**.

2.  **Rename the File**\
    Rename the `.qmd` file from `hw01.qmd` to: `hw01_yourusername.qmd`.\

For example: `hw01_semiyari.qmd`

### Working on Your Homework

-   As you complete each part of the homework:

    -   **Stage** and **commit** your `.qmd` file *after each question/sub-question*, such as `1.1`, `1.2`, etc.\
    -   **Render** your `.qmd` file to **HTML only** (no PDF or Word formats).\
    -   **Push** your updates to GitHub after completing each **main question**, such as `1.0`, `2.0`, etc.

> Committing and pushing to GitHub **are graded components** of this assignment.

## Final Checks Before Submission

-   **Re-render** to HTML before your final commit and push.\

-   **Check your output:**\

    -   Make sure **all images display correctly**.\
    -   Avoid including **excessive output or large datasets** unless asked.

## Final Submission

Once you have finished and pushed your final version:

-   Copy the link of your **GitHub Pages site** (where the rendered HTML is published).
-   **Post the link** to the Canvas assignment submission page.

```{r}
# Install pacman if it's not already installed
if (!require("pacman")) install.packages("pacman")

# Use pacman to install (if needed) and load the packages
pacman::p_load(pacman, tidyverse,
               nycflights13)
```

-   Consider the data in the nycflights13 package

## **Problem 2**: Is there a relationship between the age of a plane and its delays?

### Cleaning and structuring data

```{r}
flights2 <- flights %>% 
  select(tailnum, dep_delay) %>% 
  group_by(tailnum) %>% 
  summarize(
    mean_delay  =  mean(dep_delay, na.rm = TRUE),
    .groups = "drop"
  )

planes2 <- planes %>% 
  select(tailnum, year)


df <- planes2 %>% 
  left_join(flights2, join_by(tailnum)) %>% 
  mutate(age = 2013 - year) %>% 
  drop_na()
```

### Analysis

-   H0: beta1 = 0

-   HA: beta1 != 0

```{r}
planes.lm <- lm(mean_delay~age, data = df)
summary(planes.lm)
```

-   With a p-value of 0.0598, there is not sufficient evidence to reject the null hypothesis that there is no association between aircraft age and mean delay.

-   We conclude that there is no association between the age of an aircraft and its mean delays.

### Plots

```{r}
plot(mean_delay~age,data=df)
abline(planes.lm, col = "red")
```

## **Problem 3**: Find the 10 days of the year that have the highest median departure delay, then select all flights from those 10 days.

```{r}
head(flights)

top10delay <- flights %>% 
  mutate(date = make_datetime(year,month, day)) %>% 
  group_by(date) %>% 
  summarize(
    median_delay  =  median(dep_delay, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  arrange(desc(median_delay)) %>% 
  head(10)

flights_on_top10delay <- flights %>% 
  mutate(date = make_datetime(year,month, day)) %>%
  left_join(top10delay, join_by(date)) %>% 
  drop_na(median_delay) %>% 
  select(flight, date, dep_delay)
```
